---
pagetitle: 'Seagrant ME-NH trawl survey <br> Objective 2: Analysis of similarity and Analysis
  of variance'
author: "Jerelle Jesse"
output:
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

```

```{r, echo = FALSE}
# Access GMRI CSS Style
library(gmRi)
gmRi::use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")
```

# Analysis of Similarity and Analysis of Variance

## Data- Biomass of top 50 species
* average across depth strata using the [NOAA IEA technical document](https://noaa-edab.github.io/tech-doc/inshoresurvdat.html#data-analysis-29)
* calculate dissimilarity matrix with Bray-Curtis distances

```{r }
#load libraries
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(indicspecies)
library(here)
library(rmarkdown)
library(gridExtra)

#bring in species matrix for analysis
trawl_data_arrange<-read.csv(here("Data/species_biomass_matrix.csv"))[-1]

#separate meta data from matrix
ME_group_data<-trawl_data_arrange[, c(1,2,3,54,55,56,57,58)]
ME_NMDS_data<-as.matrix(trawl_data_arrange[,4:53])

#calculate dissimilarity matrix for tests
trawl_dist<-vegdist(ME_NMDS_data,distance="bray")

paged_table(head(trawl_data_arrange))

```

### Analysis of similarity (Anosim)

* tests statistically whether there is a significant difference between two or more groups
* works by testing if distances between groups are greater than within groups
* significant values mean that there is a statistically significant difference in the communities between the groups
* R statistic closer to 1 is more dissimilar

#### Region
```{r }
#region
ano_region<- anosim(trawl_dist, trawl_data_arrange$Region, permutations = 999)
ano_region #regions are statistically different communities 
summary(ano_region)
plot(ano_region) #regions don't look very different in plot though...confidence bands all overlap



```

#### Region grouped
```{r }
#region
ano_region_groups<- anosim(trawl_dist, trawl_data_arrange$REGION_NEW, permutations = 999)
ano_region_groups #regions are statistically different communities 
summary(ano_region_groups)
plot(ano_region_groups) #

```

#### Year
```{r}
#Time
ano_year<- anosim(trawl_dist, trawl_data_arrange$Year, permutations = 999)
ano_year #years are statistically different communities
summary(ano_year)
plot(ano_year)


```


#### Year blocks
```{r}
#Year blocks
ano_year_blocks<- anosim(trawl_dist, trawl_data_arrange$YEAR_GROUPS, permutations = 999)
ano_year_blocks #years are statistically different communities
summary(ano_year_blocks)
plot(ano_year_blocks)
```


### Analysis of variance (Adonis)

* Permanova
* tests whether there is a difference between means of groups
* works by calculating the sum of squares from the centroid of the group

#### Region
```{r}
adonis<-adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)


```

#### Year
```{r}
adonis<-adonis2(trawl_dist~Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Region and Year
```{r}
adonis<-adonis2(trawl_dist~Region*Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

mod<-betadisper(trawl_dist, ME_group_data$YEAR_GROUPS)
mod

centroids<-data.frame(grps=rownames(mod$centroids),data.frame(mod$centroids))
vectors<-data.frame(group=mod$group,data.frame(mod$vectors))
rownames(centroids)<-NULL

# to create the lines from the centroids to each point we will put it in a format that ggplot can handle
seg.data<-cbind(vectors[,1:3],centroids[rep(1:nrow(centroids),as.data.frame(table(vectors$group))$Freq),2:3])

names(seg.data)<-c("group","v.PCoA1","v.PCoA2","PCoA1","PCoA2")

# create the convex hulls of the outermost points
grp1.hull<-seg.data[seg.data$group=="2000-2004",1:3][chull(seg.data[seg.data$group=="2000-2004",2:3]),]

grp2.hull<-seg.data[seg.data$group=="2005-2009",1:3][chull(seg.data[seg.data$group=="2005-2009",2:3]),]

grp3.hull<-seg.data[seg.data$group=="2010-2014",1:3][chull(seg.data[seg.data$group=="2010-2014",2:3]),]

grp4.hull<-seg.data[seg.data$group=="2015-2019",1:3][chull(seg.data[seg.data$group=="2015-2019",2:3]),]

all.hull<-rbind(grp1.hull,grp2.hull,grp3.hull, grp4.hull)

#plot not working to show if statistical differences are in center or variation of data
# panel.a<-ggplot() +
#   geom_polygon(data=all.hull[all.hull=="2000-2004",],aes(x=v.PCoA1,y=v.PCoA2),colour="black",alpha=0,linetype="dashed") +
#   geom_segment(data=subset(seg.data, group="2000-2004"),aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),alpha=0.30) + 
#   geom_point(data=centroids[1,1:3], aes(x=PCoA1,y=PCoA2),size=4,colour="red",shape=16) + 
#   geom_point(data=subset(seg.data, group="2000-2004"), aes(x=v.PCoA1,y=v.PCoA2),size=2,shape=16) +
#   labs(title="2000-2004",x="",y="") +
#   coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5)) +
#   theme(legend.position="none")
# 
# panel.b<-ggplot() + 
#   geom_polygon(data=all.hull[all.hull=="2005-2009",],aes(x=v.PCoA1,y=v.PCoA2),colour="black",alpha=0,linetype="dashed") +
#   geom_segment(data=seg.data[31:60,],aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),alpha=0.30) + 
#   geom_point(data=centroids[2,1:3], aes(x=PCoA1,y=PCoA2),size=4,colour="red",shape=17) + 
#   geom_point(data=seg.data[31:60,], aes(x=v.PCoA1,y=v.PCoA2),size=2,shape=17) +
#   labs(title="2005-2009",x="",y="") +
#   coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5)) +
#   theme(legend.position="none")
# 
# panel.c<-ggplot() + 
#   geom_polygon(data=all.hull[all.hull=="2010-2014",],aes(x=v.PCoA1,y=v.PCoA2),colour="black",alpha=0,linetype="dashed") +
#   geom_segment(data=seg.data[61:90,],aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),alpha=0.30) +
#   geom_point(data=centroids[3,1:3], aes(x=PCoA1,y=PCoA2),size=4,colour="red",shape=15) + 
#   geom_point(data=seg.data[61:90,], aes(x=v.PCoA1,y=v.PCoA2),size=2,shape=15) + 
#   labs(title="2010-2014",x="",y="") +
#   coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5)) +
#   theme(legend.position="none")
# 
# panel.d<-ggplot() + 
#   geom_polygon(data=all.hull[all.hull=="2015-2019",],aes(x=v.PCoA1,y=v.PCoA2),colour="black",alpha=0,linetype="dashed") +
#   geom_segment(data=seg.data[61:90,],aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),alpha=0.30) +
#   geom_point(data=centroids[4,1:3], aes(x=PCoA1,y=PCoA2),size=4,colour="red",shape=12) + 
#   geom_point(data=seg.data[61:90,], aes(x=v.PCoA1,y=v.PCoA2),size=2,shape=15) + 
#   labs(title="2015-2019",x="",y="") +
#   coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5)) +
#   theme(legend.position="none")
# 
# grid.arrange(panel.a,panel.b,panel.c,panel.d,nrow=1)

```

#### Region and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~Region*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

#### Region groups
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~REGION_NEW, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```


#### Region groups and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~REGION_NEW*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

### Pairwise

* Vegan does not have a function for this, but I found a wrapper that seems frequently used on github
* select groups to test, one pair at a time 
* Adjust p-values for multiple tests

#### Region
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair 

```

#### Region groups
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~REGION_NEW, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair 

```

#### Year blocks
```{r}
#pair-wise test to see what is different for year blocks
pair<-pairwise.adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair
```

### Dispersion 

* anosim very sensitive to heterogeneity (Anderson and Walsh 2013)
* Could get false significant results from differences in variance instead of mean
* adonis is less affected by heterogeneity for *balanced designs* 
* PRIMER can deal with dispersion issues, but vegan does not yet
* tests null hypothesis that there is no difference in dispersion between groups
* p-value <0.05 means difference is significant

#### Region
```{r}
#betadisper test homogeneity of dispersion among groups
#Region
bd<-betadisper(trawl_dist,ME_group_data$Region) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)  
boxplot(bd)

```

#### Region group
```{r}
#Region
bd<-betadisper(trawl_dist,ME_group_data$REGION_NEW) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)  
boxplot(bd)
```

#### Year
```{r}
#Year
bd<-betadisper(trawl_dist,ME_group_data$Year) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)
boxplot(bd)

```


#### Year block
```{r}
#Year blocks
bd<-betadisper(trawl_dist,ME_group_data$YEAR_GROUPS) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)
boxplot(bd)
```

### Indicator species analysis

* test if a species if found significantly more in one group compared to another
* all combinations of groups

#### Region
```{r}
#see which species are found significantly more in each Region
inv_region<-multipatt(ME_NMDS_data, ME_group_data$Region, func = "r.g", control = how(nperm=999))
summary(inv_region)

```

#### Year block
```{r}
#see which species are found significantly more in each Region
inv_year<-multipatt(ME_NMDS_data, ME_group_data$YEAR_GROUPS, func = "r.g", control = how(nperm=999))
summary(inv_year)
```


## Data- Abundance of top 50 species
* average across depth strata using the [NOAA IEA technical document](https://noaa-edab.github.io/tech-doc/inshoresurvdat.html#data-analysis-29)
* calculate dissimilarity matrix with Bray-Curtis distances

```{r echo=FALSE, message=FALSE}
#load libraries
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(indicspecies)
library(here)
library(rmarkdown)

#bring in species matrix for analysis
trawl_data_arrange<-read.csv(here("Data/species_abundance_matrix.csv"))[-1]

#separate meta data from matrix
ME_group_data<-trawl_data_arrange[, c(1,2,3,54,55,56,57,58)]
ME_NMDS_data<-as.matrix(trawl_data_arrange[,4:53])

#calculate dissimilarity matrix for tests
trawl_dist<-vegdist(ME_NMDS_data,distance="bray")

paged_table(head(trawl_data_arrange))

```

### Analysis of similarity (Anosim)

* tests statistically whether there is a significant difference between two or more groups
* works by testing if distances between groups are greater than within groups
* significant values mean that there is a statistically significant difference in the communities between the groups
* R statistic closer to 1 is more dissimilar

#### Region
```{r }
#region
ano_region<- anosim(trawl_dist, trawl_data_arrange$Region, permutations = 999)
ano_region #regions are statistically different communities 
summary(ano_region)
plot(ano_region) #regions don't look very different in plot though...confidence bands all overlap



```

#### Region grouped
```{r }
#region
ano_region_groups<- anosim(trawl_dist, trawl_data_arrange$REGION_NEW, permutations = 999)
ano_region_groups #regions are statistically different communities 
summary(ano_region_groups)
plot(ano_region_groups) #

```

#### Year
```{r}
#Time
ano_year<- anosim(trawl_dist, trawl_data_arrange$Year, permutations = 999)
ano_year #years are statistically different communities
summary(ano_year)
plot(ano_year)


```


#### Year blocks
```{r}
#Year blocks
ano_year_blocks<- anosim(trawl_dist, trawl_data_arrange$YEAR_GROUPS, permutations = 999)
ano_year_blocks #years are statistically different communities
summary(ano_year_blocks)
plot(ano_year_blocks)
```


### Analysis of variance (Adonis)

* Permanova
* tests whether there is a difference between means of groups
* works by calculating the sum of squares from the centroid of the group

#### Region
```{r}
adonis<-adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Year
```{r}
adonis<-adonis2(trawl_dist~Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Region and Year
```{r}
adonis<-adonis2(trawl_dist~Region*Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

#### Region and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~Region*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

#### Region groups 
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~REGION_NEW, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

#### Region groups and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~REGION_NEW*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

### Pairwise

* Vegan does not have a function for this, but I found a wrapper that seems frequently used on github
* select groups to test, one pair at a time 
* Adjust p-values for multiple tests

#### Region
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair #shows all the regions are significantly different except 3 and 4

```

#### Region groups
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~REGION_NEW, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair #shows all the regions are significantly different except 3 and 4

```

#### Year blocks
```{r}
#pair-wise test to see what is different for year blocks
pair<-pairwise.adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair
```

### Dispersion 

* anosim very sensitive to heterogeneity (Anderson and Walsh 2013)
* Could get false significant results from differences in variance instead of mean
* adonis is less affected by heterogeneity for *balanced designs* 
* PRIMER can deal with dispersion issues, but vegan does not yet
* tests null hypothesis that there is no difference in dispersion between groups
* p-value <0.05 means difference is significant

#### Region
```{r}
#betadisper test homogeneity of dispersion among groups
#Region
bd<-betadisper(trawl_dist,ME_group_data$Region) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)  
boxplot(bd)
```

#### Region group
```{r}
#Region
bd<-betadisper(trawl_dist,ME_group_data$REGION_NEW) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)  
boxplot(bd)
```

#### Year
```{r}
#Year
bd<-betadisper(trawl_dist,ME_group_data$Year) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)
boxplot(bd)

```


#### Year block
```{r}
#Year blocks
bd<-betadisper(trawl_dist,ME_group_data$YEAR_GROUPS) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)
boxplot(bd)

```

### Indicator species analysis

* test if a species if found significantly more in one group compared to another
* all combinations of groups

#### Region
```{r}
#see which species are found significantly more in each Region
inv_region<-multipatt(ME_NMDS_data, ME_group_data$Region, func = "r.g", control = how(nperm=999))
summary(inv_region)

```

#### Year block
```{r}
#see which species are found significantly more in each Region
inv_year<-multipatt(ME_NMDS_data, ME_group_data$YEAR_GROUPS, func = "r.g", control = how(nperm=999))
summary(inv_year)
```



## Data- Biomass of functional groups
* average across depth strata using the [NOAA IEA technical document](https://noaa-edab.github.io/tech-doc/inshoresurvdat.html#data-analysis-29)
* calculate dissimilarity matrix with Bray-Curtis distances

```{r echo=FALSE, message=FALSE}
#load libraries
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(indicspecies)
library(here)
library(rmarkdown)

#bring in species matrix for analysis
trawl_data_arrange<-read.csv(here("Data/group_biomass_matrix.csv"))[-1]

#separate meta data from matrix
ME_group_data<-trawl_data_arrange[, c(1,2,3,9,10,11,12,13)]
ME_NMDS_data<-as.matrix(trawl_data_arrange[,4:8])

#calculate dissimilarity matrix for tests
trawl_dist<-vegdist(ME_NMDS_data,distance="bray")

paged_table(head(trawl_data_arrange))

```

### Analysis of similarity (Anosim)

* tests statistically whether there is a significant difference between two or more groups
* works by testing if distances between groups are greater than within groups
* significant values mean that there is a statistically significant difference in the communities between the groups
* R statistic closer to 1 is more dissimilar

#### Region
```{r }
#region
ano_region<- anosim(trawl_dist, trawl_data_arrange$Region, permutations = 999)
ano_region #regions are statistically different communities 
summary(ano_region)
plot(ano_region) #regions don't look very different in plot though...confidence bands all overlap



```

#### Region grouped
```{r }
#region
ano_region_groups<- anosim(trawl_dist, trawl_data_arrange$REGION_NEW, permutations = 999)
ano_region_groups #regions are statistically different communities 
summary(ano_region_groups)
plot(ano_region_groups) #

```

#### Year
```{r}
#Time
ano_year<- anosim(trawl_dist, trawl_data_arrange$Year, permutations = 999)
ano_year #years are statistically different communities
summary(ano_year)
plot(ano_year)


```


#### Year blocks
```{r}
#Year blocks
ano_year_blocks<- anosim(trawl_dist, trawl_data_arrange$YEAR_GROUPS, permutations = 999)
ano_year_blocks #years are statistically different communities
summary(ano_year_blocks)
plot(ano_year_blocks)
```


### Analysis of variance (Adonis)

* Permanova
* tests whether there is a difference between means of groups
* works by calculating the sum of squares from the centroid of the group

#### Region
```{r}
adonis<-adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Year
```{r}
adonis<-adonis2(trawl_dist~Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Region and Year
```{r}
adonis<-adonis2(trawl_dist~Region*Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

#### Region and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~Region*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```


#### Region groups 
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~REGION_NEW, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

#### Region groups and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~REGION_NEW*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

### Pairwise

* Vegan does not have a function for this, but I found a wrapper that seems frequently used on github
* select groups to test, one pair at a time 
* Adjust p-values for multiple tests

#### Region
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair 

```

#### Region groups
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~REGION_NEW, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair 

```

#### Year blocks
```{r}
#pair-wise test to see what is different for year blocks
pair<-pairwise.adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair
```

### Dispersion 

* anosim very sensitive to heterogeneity (Anderson and Walsh 2013)
* Could get false significant results from differences in variance instead of mean
* adonis is less affected by heterogeneity for *balanced designs* 
* PRIMER can deal with dispersion issues, but vegan does not yet
* tests null hypothesis that there is no difference in dispersion between groups
* p-value <0.05 means difference is significant

#### Region
```{r}
#betadisper test homogeneity of dispersion among groups
#Region
bd<-betadisper(trawl_dist,ME_group_data$Region) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)  
boxplot(bd)

```

#### Region group
```{r}
#Region
bd<-betadisper(trawl_dist,ME_group_data$REGION_NEW) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)  
boxplot(bd)
```

#### Year
```{r}
#Year
bd<-betadisper(trawl_dist,ME_group_data$Year) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)
boxplot(bd)

```


#### Year block
```{r}
#Year blocks
bd<-betadisper(trawl_dist,ME_group_data$YEAR_GROUPS) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)
boxplot(bd)

```

### Indicator species analysis

* test if a species if found significantly more in one group compared to another
* all combinations of groups

#### Region
```{r}
#see which species are found significantly more in each Region
inv_region<-multipatt(ME_NMDS_data, ME_group_data$Region, func = "r.g", control = how(nperm=999))
summary(inv_region)

```

#### Year block
```{r}
#see which species are found significantly more in each Region
inv_year<-multipatt(ME_NMDS_data, ME_group_data$YEAR_GROUPS, func = "r.g", control = how(nperm=999))
summary(inv_year)
```

## Data- Abundance functional groups
* average across depth strata using the [NOAA IEA technical document](https://noaa-edab.github.io/tech-doc/inshoresurvdat.html#data-analysis-29)
* calculate dissimilarity matrix with Bray-Curtis distances

```{r echo=FALSE, message=FALSE}
#load libraries
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(indicspecies)
library(here)
library(rmarkdown)

#bring in species matrix for analysis
trawl_data_arrange<-read.csv(here("Data/group_abundance_matrix.csv"))[-1]
trawl_data_arrange<-filter(trawl_data_arrange, !is.na(benthos))

#separate meta data from matrix
ME_group_data<-trawl_data_arrange[, c(1,2,3,9,10,11,12,13)]
ME_NMDS_data<-as.matrix(trawl_data_arrange[,4:8])

#calculate dissimilarity matrix for tests
trawl_dist<-vegdist(ME_NMDS_data,distance="bray")

paged_table(head(trawl_data_arrange))

```

### Analysis of similarity (Anosim)

* tests statistically whether there is a significant difference between two or more groups
* works by testing if distances between groups are greater than within groups
* significant values mean that there is a statistically significant difference in the communities between the groups
* R statistic closer to 1 is more dissimilar

#### Region
```{r }
#region
ano_region<- anosim(trawl_dist, trawl_data_arrange$Region, permutations = 999)
ano_region #regions are statistically different communities 
summary(ano_region)
plot(ano_region) #regions don't look very different in plot though...confidence bands all overlap



```

#### Region grouped
```{r }
#region
ano_region_groups<- anosim(trawl_dist, trawl_data_arrange$REGION_NEW, permutations = 999)
ano_region_groups #regions are statistically different communities 
summary(ano_region_groups)
plot(ano_region_groups) #

```

#### Year
```{r}
#Time
ano_year<- anosim(trawl_dist, trawl_data_arrange$Year, permutations = 999)
ano_year #years are statistically different communities
summary(ano_year)
plot(ano_year)


```


#### Year blocks
```{r}
#Year blocks
ano_year_blocks<- anosim(trawl_dist, trawl_data_arrange$YEAR_GROUPS, permutations = 999)
ano_year_blocks #years are statistically different communities
summary(ano_year_blocks)
plot(ano_year_blocks)
```


### Analysis of variance (Adonis)

* Permanova
* tests whether there is a difference between means of groups
* works by calculating the sum of squares from the centroid of the group

#### Region and Year
```{r}
adonis<-adonis2(trawl_dist~Region*Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Region 
```{r}
adonis<-adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Year
```{r}
adonis<-adonis2(trawl_dist~Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

#### Year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

#### Region and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~Region*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```


#### Region groups 
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~REGION_NEW, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

#### Region groups and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~REGION_NEW*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

### Pairwise

* Vegan does not have a function for this, but I found a wrapper that seems frequently used on github
* select groups to test, one pair at a time 
* Adjust p-values for multiple tests

#### Region
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair 

```

#### Region groups
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~REGION_NEW, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair 

```

#### Year blocks
```{r}
#pair-wise test to see what is different for year blocks
pair<-pairwise.adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair
```

### Dispersion 

* anosim very sensitive to heterogeneity (Anderson and Walsh 2013)
* Could get false significant results from differences in variance instead of mean
* adonis is less affected by heterogeneity for *balanced designs* 
* PRIMER can deal with dispersion issues, but vegan does not yet
* tests null hypothesis that there is no difference in dispersion between groups
* p-value <0.05 means difference is significant

#### Region
```{r}
#betadisper test homogeneity of dispersion among groups
#Region
bd<-betadisper(trawl_dist,ME_group_data$Region) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)  
boxplot(bd)

```

#### Region group
```{r}
#Region
bd<-betadisper(trawl_dist,ME_group_data$REGION_NEW) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)  
boxplot(bd)

```

#### Year
```{r}
#Year
bd<-betadisper(trawl_dist,ME_group_data$Year) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)
boxplot(bd)

```


#### Year block
```{r}
#Year blocks
bd<-betadisper(trawl_dist,ME_group_data$YEAR_GROUPS) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd, hull=FALSE, ellipse = TRUE)
boxplot(bd)

```

### Indicator species analysis

* test if a species if found significantly more in one group compared to another
* all combinations of groups

#### Region
```{r}
#see which species are found significantly more in each Region
inv_region<-multipatt(ME_NMDS_data, ME_group_data$Region, func = "r.g", control = how(nperm=999))
summary(inv_region)

```

#### Year block
```{r}
#see which species are found significantly more in each Region
inv_year<-multipatt(ME_NMDS_data, ME_group_data$YEAR_GROUPS, func = "r.g", control = how(nperm=999))
summary(inv_year)
```

