---
pagetitle: 'Seagrant ME-NH trawl survey <br> Objective 2: Analysis of similarity and Analysis
  of variance'
author: "Jerelle Jesse"
output:
  html_document: 
    toc: true
    toc_float: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of Similarity and Analysis of Variance

## Data

* Biomass of top 50 species
* average across depth strata using the [NOAA IEA technical document](https://noaa-edab.github.io/tech-doc/inshoresurvdat.html#data-analysis-29)
* calculate dissimilarity matrix with Bray-Curtis distances

```{r echo=FALSE, message=FALSE}
#load libraries
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(indicspecies)
library(here)
library(rmarkdown)

#bring in species matrix for analysis
trawl_data_arrange<-read.csv(here("Data/species_biomass_matrix.csv"))[-1]

#separate meta data from matrix
ME_group_data<-trawl_data_arrange[, c(1,2,3,54,55,56,57,58)]
ME_NMDS_data<-as.matrix(trawl_data_arrange[,4:53])

#calculate dissimilarity matrix for tests
trawl_dist<-vegdist(ME_NMDS_data,distance="bray")

paged_table(head(trawl_data_arrange))

```

## Analysis of similarity (Anosim)

* tests statistically whether there is a significant difference between two or more groups
* works by testing if distances between groups are greater than within groups
* significant values mean that there is a statistically significant difference in the communities between the groups
* R statistic closer to 1 is more dissimilar

### Region
```{r warning=FALSE}
#region
ano_region<- anosim(trawl_dist, trawl_data_arrange$Region, permutations = 999)
ano_region #regions are statistically different communities 
summary(ano_region)
plot(ano_region) #regions don't look very different in plot though...confidence bands all overlap



```

### Year
```{r}
#Time
ano_year<- anosim(trawl_dist, trawl_data_arrange$Year, permutations = 999)
ano_year #years are statistically different communities
summary(ano_year)
plot(ano_year)


```


### Year blocks
```{r}
#Year blocks
ano_year_blocks<- anosim(trawl_dist, trawl_data_arrange$YEAR_GROUPS, permutations = 999)
ano_year_blocks #years are statistically different communities
summary(ano_year_blocks)
plot(ano_year_blocks)
```


## Analysis of variance (Adonis)

* Permanova
* tests whether there is a difference between means of groups
* works by calculating the sum of squares from the centroid of the group

### Region and Year
```{r}
adonis<-adonis2(trawl_dist~Region*Year, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)

```

### Region and year block
```{r}
#with year blocks
adonis<-adonis2(trawl_dist~Region*YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
adonis
summary(adonis)
```

## Pairwise

* Vegan does not have a function for this, but I found a wrapper that seems frequently used on github
* select groups to test, one pair at a time 
* Adjust p-values for multiple tests

### Region
```{r}
#pair-wise test to see what is different
pair<-pairwise.adonis2(trawl_dist~Region, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair #shows all the regions are significantly different except 3 and 4

```

### Year blocks
```{r}
#pair-wise test to see what is different for year blocks
pair<-pairwise.adonis2(trawl_dist~YEAR_GROUPS, data=ME_group_data, by="terms", permutations = 9999)
summary(pair)
pair
```

## Dispersion 

* anosim very sensitive to heterogeneity (Anderson and Walsh 2013)
* Could get false significant results from differences in variance instead of mean
* adonis is less affected by heterogeneity for *balanced designs* 
* PRIMER can deal with dispersion issues, but vegan does not yet
* tests null hypothesis that there is no difference in dispersion between groups
* p-value <0.05 means difference is significant

### Region
```{r}
#betadisper test homogeneity of dispersion among groups
#Region
bd<-betadisper(trawl_dist,ME_group_data$Region) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd)  

```

### Year
```{r}
#Year
bd<-betadisper(trawl_dist,ME_group_data$Year) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd)


```


### Year block
```{r}
#Year blocks
bd<-betadisper(trawl_dist,ME_group_data$YEAR_GROUPS) 
bd
anova(bd) 
#test based on permutations
permutest(bd)

plot(bd)
```

## Indicator species analysis

* test if a species if found significantly more in one group compared to another
* all combinations of groups

### Region
```{r}
#see which species are found significantly more in each Region
inv_region<-multipatt(ME_NMDS_data, ME_group_data$Region, func = "r.g", control = how(nperm=999))
summary(inv_region)

```

### Year block
```{r}
#see which species are found significantly more in each Region
inv_year<-multipatt(ME_NMDS_data, ME_group_data$YEAR_GROUPS, func = "r.g", control = how(nperm=999))
summary(inv_year)
```

