---
pagetitle: 'Seagrant ME-NH trawl survey <br> Objective 1: Generalized Additive Mixed Models'
author: "Jerelle Jesse"
output:
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

```

```{r, echo = FALSE}
# Access GMRI CSS Style
library(gmRi)
gmRi::use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")
```

## Generalized additive mixed models

```{r message=FALSE, warning=FALSE, include=FALSE}

library(mgcv)
library(MASS)
library(gamm4)

```


## Fall ME-NH GAMMs
*  landings data and FVCOM already added to survey data in previous code

```{r message=FALSE, warning=FALSE, include=FALSE}
# trawl observations 
setwd("C:/Users/jjesse/Box/Kerr Lab/Fisheries Science Lab/ME NH Trawl- Seagrant/Seagrant-AEW/Results/GAMMs")
fall <- read.csv("ME_NH_fall_exp.csv")

# FVCOM observations 
fall_fvcom <- read.csv("ME_NH_fall_full.csv")
```

### Species richness
#### with survey environmental data
```{r}

#1 choose response distribution - start w/normal distribution
hist(fall$N_species) # start w/normal distribution

#2 choose k - let GCV find optimal 

#3 autocorrelation? 
# lat/long = correlated
# bottom/surface salinity = correlated
#plot(fall[,20], fall[,23])
# yes so fit w/GAMM

#4 is k large enough? diagnostics ok?
# diagnostic/residual plots; QQ,resid vs. pred
# take care when interpretting results
# k-index; further below 1 = missed pattern in resids
# k is too low if edf ~ k'

## best model fit is N_Fall_2
N_Fall_2 <- gamm4(N_species ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(metric_tons) + s(SURFACE_SALINITY) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall) 

gam.check(N_Fall_2$gam)
# plot(resid(N_Fall_2$gam))
# abline(h = 0)
#mean(resid(N_Fall_2$gam)^2)


#5 significant trend? 
# interpretting results
summary(N_Fall_2$gam) # importance of terms 
print(N_Fall_2$gam) # edf; higher = more complex splines 
#confint(N_Fall_2$gam)

plot(N_Fall_2$gam)

```


#### with FVCOM data
```{r}
N_Fall_FV <- gamm4(N_species ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall_fvcom) 

gam.check(N_Fall_FV$gam)
summary(N_Fall_FV$gam)
#plot(resid(N_Fall_FV$gam))
plot(N_Fall_FV$gam)

```

### Shannon-Weiner Diversity
#### survey environmental data
```{r}
H_Fall <-  gamm4(H_index ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) + s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) + s(metric_tons), random = ~ (1|YEAR), data = fall) 

gam.check(H_Fall$gam)
summary(H_Fall$gam)
plot(H_Fall$gam)

```

#### FVCOM data
```{r}
H_Fall_FV <- gamm4(H_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall_fvcom) 

gam.check(H_Fall_FV$gam)
summary(H_Fall_FV$gam)
plot(H_Fall_FV$gam)

```

### Simpson's Diversity
#### survey environmental data

```{r}
D_Fall <- gamm4(D_index  ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +
                  s(SURFACE_SALINITY)  + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) + s(metric_tons), random = ~ (1|YEAR), data = fall)

gam.check(D_Fall$gam)
summary(D_Fall$gam)
plot(D_Fall$gam)
```

#### FVCOM data
```{r}
D_Fall_FV <- gamm4(D_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) +s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall_fvcom) 

gam.check(D_Fall_FV$gam)
summary(D_Fall_FV$gam)
plot(D_Fall_FV$gam)
```


### Simpson's Evenness
#### survey environmental data

```{r}
E_Fall <- gamm4(E_index  ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) + s(metric_tons), random = ~ (1|YEAR), data = fall) 

gam.check(E_Fall$gam)
summary(E_Fall$gam)
plot(E_Fall$gam)

```

#### FVCOM data
```{r}
E_Fall_FV <- gamm4(E_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall_fvcom) 

gam.check(E_Fall_FV$gam)
summary(E_Fall_FV$gam)
plot(E_Fall_FV$gam)

```


### Taxonomic diversity
#### survey environmental data

```{r}
delta_Fall <- gamm4(delta ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) + s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) +  s(metric_tons), random = ~ (1|YEAR), data = fall)

gam.check(delta_Fall$gam)
summary(delta_Fall$gam)
plot(delta_Fall$gam)
```

#### FVCOM data
```{r}
delta_Fall_FV <- gamm4(delta ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) +s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall_fvcom) 

gam.check(delta_Fall_FV$gam)
summary(delta_Fall_FV$gam)
plot(delta_Fall_FV$gam)
```


### Taxonomic distinctness
#### survey environmental data

```{r}
delta_star_Fall <- gamm4(delta_star ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) +  s(metric_tons), random = ~ (1|YEAR), data = fall)

gam.check(delta_star_Fall$gam)
summary(delta_star_Fall$gam)
plot(delta_star_Fall$gam)
```

#### FVCOM data
```{r}
delta_star_Fall_FV <- gamm4(delta_star ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall_fvcom)

gam.check(delta_star_Fall_FV$gam)
summary(delta_star_Fall_FV$gam)
plot(delta_star_Fall_FV$gam)
```


### Average taxonomic distinctness
#### survey environmental data

```{r}
delta_plus_Fall <- gamm4(delta_plus ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) +  s(metric_tons), random = ~ (1|YEAR), data = fall_fvcom)

gam.check(delta_plus_Fall$gam)
summary(delta_plus_Fall$gam)
plot(delta_plus_Fall$gam)
```

#### FVCOM data
```{r}
delta_plus_Fall_FV <- gamm4(delta_plus ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall_fvcom)

gam.check(delta_plus_Fall_FV$gam)
summary(delta_plus_Fall_FV$gam)
plot(delta_plus_Fall_FV$gam)
```


### Variation in taxonomic distinctness
#### survey environmental data

```{r}
delta_var_Fall <- gamm4(delta_var ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) +  s(metric_tons), random = ~ (1|YEAR), data = fall)

gam.check(delta_var_Fall$gam)
summary(delta_var_Fall$gam)
plot(delta_var_Fall$gam)
```

#### FVCOM data
```{r}
delta_var_Fall_FV <- gamm4(delta_var ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) +   s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = fall_fvcom)

gam.check(delta_var_Fall_FV$gam)
summary(delta_var_Fall_FV$gam)
plot(delta_var_Fall_FV$gam)
```


## Spring ME-NH GAMMs
*  landings data and FVCOM already added to survey data in previous code

```{r}
# trawl observations 
setwd("C:/Users/jjesse/Box/Kerr Lab/Fisheries Science Lab/ME NH Trawl- Seagrant/Seagrant-AEW/Results/GAMMs")

# trawl observations:
spring <- read.csv("ME_NH_spring_exp.csv")

# FVCOM observations 
spring_fvcom <- read.csv("ME_NH_spring_full.csv")

```

### Species richness
#### with survey environmental data
```{r}

#1 choose response distribution - start w/normal distribution
hist(spring$N_species) # start w/normal distribution

#2 choose k - let GCV find optimal 

#3 autocorrelation? 
# lat/long = correlated
# bottom/surface salinity = correlated
#plot(spring[,20], spring[,23])
# yes so fit w/GAMM

#4 is k large enough? diagnostics ok?
# diagnostic/residual plots; QQ,resid vs. pred
# take care when interpretting results
# k-index; further below 1 = missed pattern in resids
# k is too low if edf ~ k'

## best model fit is N_Fall_2
N_spring_2 <- gamm4(N_species ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(metric_tons) + s(SURFACE_SALINITY) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring) 

gam.check(N_spring_2$gam)
#plot(resid(N_spring_2$gam))
#abline(h = 0)
#mean(resid(N_spring_2$gam)^2)


#5 significant trend? 
# interpretting results
summary(N_spring_2$gam) # importance of terms 
print(N_spring_2$gam) # edf; higher = more complex splines 
#confint(N_spring_2$gam)

plot(N_spring_2$gam)

```


#### with FVCOM data
```{r}
N_spring_FV <- gamm4(N_species ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring_fvcom) 

gam.check(N_spring_FV$gam)
summary(N_spring_FV$gam)
plot(resid(N_spring_FV$gam))
plot(N_spring_FV$gam)

```

### Shannon-Weiner Diversity
#### survey environmental data
```{r}
H_spring <-  gamm4(H_index ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) + s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) + s(metric_tons), random = ~ (1|YEAR), data = spring) 

gam.check(H_spring$gam)
summary(H_spring$gam)
plot(H_spring$gam)

```

#### FVCOM data
```{r}
H_spring_FV <- gamm4(H_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring_fvcom) 

gam.check(H_spring_FV$gam)
summary(H_spring_FV$gam)
plot(H_spring_FV$gam)

```

### Simpson's Diversity
#### survey environmental data

```{r}
D_spring <- gamm4(D_index  ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) + s(SURFACE_SALINITY)  + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) + s(metric_tons), random = ~ (1|YEAR), data = spring)

gam.check(D_spring$gam)
summary(D_spring$gam)
plot(D_spring$gam)
```

#### FVCOM data
```{r}
D_spring_FV <- gamm4(D_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) +s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring_fvcom) 

gam.check(D_spring_FV$gam)
summary(D_spring_FV$gam)
plot(D_spring_FV$gam)
```


### Simpson's Evenness
#### survey environmental data

```{r}
E_spring <- gamm4(E_index  ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) + s(metric_tons), random = ~ (1|YEAR), data = spring) 

gam.check(E_spring$gam)
summary(E_spring$gam)
plot(E_spring$gam)

```

#### FVCOM data
```{r}
E_spring_FV <- gamm4(E_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring_fvcom) 

gam.check(E_spring_FV$gam)
summary(E_spring_FV$gam)
plot(E_spring_FV$gam)

```


### Taxonomic diversity
#### survey environmental data

```{r}
delta_spring <- gamm4(delta ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) + s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) +  s(metric_tons), random = ~ (1|YEAR), data = spring)

gam.check(delta_spring$gam)
summary(delta_spring$gam)
plot(delta_spring$gam)
```

#### FVCOM data
```{r}
delta_spring_FV <- gamm4(delta ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) +s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring_fvcom) 

gam.check(delta_spring_FV$gam)
summary(delta_spring_FV$gam)
plot(delta_spring_FV$gam)
```


### Taxonomic distinctness
#### survey environmental data

```{r}
delta_star_spring <- gamm4(delta_star ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) +  s(metric_tons), random = ~ (1|YEAR), data = spring)

gam.check(delta_star_spring$gam)
summary(delta_star_spring$gam)
plot(delta_star_spring$gam)
```

#### FVCOM data
```{r}
delta_star_spring_FV <- gamm4(delta_star ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring_fvcom)

gam.check(delta_star_spring_FV$gam)
summary(delta_star_spring_FV$gam)
plot(delta_star_spring_FV$gam)
```


### Average taxonomic distinctness
#### survey environmental data

```{r}
delta_plus_spring <- gamm4(delta_plus ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) +  s(metric_tons), random = ~ (1|YEAR), data = spring_fvcom)

gam.check(delta_plus_spring$gam)
summary(delta_plus_spring$gam)
plot(delta_plus_spring$gam)
```

#### FVCOM data
```{r}
delta_plus_spring_FV <- gamm4(delta_plus ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) + s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring_fvcom)

gam.check(delta_plus_spring_FV$gam)
summary(delta_plus_spring_FV$gam)
plot(delta_plus_spring_FV$gam)
```


### Variation in taxonomic distinctness
#### survey environmental data

```{r}
delta_var_spring <- gamm4(delta_var ~ s(WATER_TEMP_C) + s(SURFACE_TEMP_C) + s(SALINITY) +  s(SURFACE_SALINITY) + s(START_DEPTH)+ s(START_LATITUDE, START_LONGITUDE) +  s(metric_tons), random = ~ (1|YEAR), data = spring)

gam.check(delta_var_spring$gam)
summary(delta_var_spring$gam)
plot(delta_var_spring$gam)
```

#### FVCOM data
```{r}
delta_var_spring_FV <- gamm4(delta_var ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(metric_tons) +   s(FV_surf_sal) +  s(START_DEPTH) + s(START_LATITUDE, START_LONGITUDE), random = ~ (1|YEAR) , data = spring_fvcom)

gam.check(delta_var_spring_FV$gam)
summary(delta_var_spring_FV$gam)
plot(delta_var_spring_FV$gam)
```


## Fall GOM GAMMs
* all with FVCOM data
```{r}
setwd("C:/Users/jjesse/Box/Kerr Lab/Fisheries Science Lab/ME NH Trawl- Seagrant/Seagrant-AEW/Results/GAMMs")

GOM_fall <- read.csv("GOM_fall_full.csv")

GOM_spring <- read.csv("GOM_spring_full.csv")
```

### Species richness
```{r}
FV_GOM_N_FL <- gamm4(N_species ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) + s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_fall) 

gam.check(FV_GOM_N_FL$gam)
summary(FV_GOM_N_FL$gam)
plot(FV_GOM_N_FL$gam)
```

### Shannon-Wiener
```{r}
FV_GOM_H_FL <- gamm4(H_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_fall) 

gam.check(FV_GOM_H_FL$gam)
summary(FV_GOM_H_FL$gam)
plot(FV_GOM_H_FL$gam)
```

### Simpson's Dversity
```{r}
FV_GOM_D_FL <- gamm4(D_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) + s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_fall) 

gam.check(FV_GOM_D_FL$gam)
summary(FV_GOM_D_FL$gam)
plot(FV_GOM_D_FL$gam)
```


### Simpson's Evenness
```{r}
FV_GOM_E_FL <- gamm4(E_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +   s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_fall) 

gam.check(FV_GOM_E_FL$gam)
summary(FV_GOM_E_FL$gam)
plot(FV_GOM_E_FL$gam)
```


### Taxonomic diversity
```{r}
FV_GOM_delta_FL <- gamm4(delta ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_fall) 

gam.check(FV_GOM_delta_FL$gam)
summary(FV_GOM_delta_FL$gam)
plot(FV_GOM_delta_FL$gam)
```


### Taxonomic distinctness
```{r}
FV_GOM_delta_star_FL <- gamm4(delta_star ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_fall) 

gam.check(FV_GOM_delta_star_FL$gam)
summary(FV_GOM_delta_star_FL$gam)
plot(FV_GOM_delta_star_FL$gam)
```


### Average taxonomic distinctness
```{r}
FV_GOM_delta_plus_FL <- gamm4(delta_plus ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_fall) 

gam.check(FV_GOM_delta_plus_FL$gam)
summary(FV_GOM_delta_plus_FL$gam)
plot(FV_GOM_delta_plus_FL$gam)
```


### Variation in taxonomic distinctness
```{r}
FV_GOM_delta_var_FL <- gamm4(delta_var ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +   s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_fall) 

gam.check(FV_GOM_delta_var_FL$gam)
summary(FV_GOM_delta_var_FL$gam)
plot(FV_GOM_delta_var_FL$gam)
```

## Spring GOM GAMMs
### Species richness
```{r}
FV_GOM_N_SP <- gamm4(N_species ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) + s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_spring) 

gam.check(FV_GOM_N_SP$gam)
summary(FV_GOM_N_SP$gam)
plot(FV_GOM_N_SP$gam)
```


### Shannon-Wiener diversity
```{r}
FV_GOM_H_SP <- gamm4(H_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_spring) 

gam.check(FV_GOM_H_SP$gam)
summary(FV_GOM_H_SP$gam)
plot(FV_GOM_H_SP$gam)
```


### Simpson's diversity
```{r}
FV_GOM_D_SP <- gamm4(D_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +   s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_spring) 

gam.check(FV_GOM_D_SP$gam)
summary(FV_GOM_D_SP$gam)
plot(FV_GOM_D_SP$gam)
```


### Simpson's evenness
```{r}
FV_GOM_E_SP <- gamm4(E_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_spring) 

gam.check(FV_GOM_E_SP$gam)
summary(FV_GOM_E_SP$gam)
plot(FV_GOM_E_SP$gam)
```


### Taxonomic diversity
```{r}

FV_GOM_delta_SP <- gamm4(delta ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_spring) 

gam.check(FV_GOM_delta_SP$gam)
summary(FV_GOM_delta_SP$gam)
plot(FV_GOM_delta_SP$gam)
```


### Taxonomic distinctness
```{r}
FV_GOM_delta_star_SP <- gamm4(delta_star ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +   s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_spring) 

gam.check(FV_GOM_delta_star_SP$gam)
summary(FV_GOM_delta_star_SP$gam)
plot(FV_GOM_delta_star_SP$gam)
```


### Average taxonomic distinctness
```{r}
FV_GOM_delta_plus_SP <- gamm4(delta_plus ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_spring) 

gam.check(FV_GOM_delta_plus_SP$gam)
summary(FV_GOM_delta_plus_SP$gam)
plot(FV_GOM_delta_plus_SP$gam)
```


### Variation in taxonomic distinctness
```{r}
FV_GOM_delta_var_SP <- gamm4(delta_var ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(SETDEPTH)+ s(DECDEG_BEGLAT, DECDEG_BEGLON) , random = ~ (1|EST_YEAR), data = GOM_spring) 

gam.check(FV_GOM_delta_var_SP$gam)
summary(FV_GOM_delta_var_SP$gam)
plot(FV_GOM_delta_var_SP$gam)
```


## Fall MADMF GAMMs
* all with FVCOM data
```{r}
setwd("C:/Users/jjesse/Box/Kerr Lab/Fisheries Science Lab/ME NH Trawl- Seagrant/Seagrant-AEW/Results/GAMMs")

MA_fall <- read.csv("MA_fall_full.csv")

MA_spring <- read.csv("MA_spring_full.csv")
```

### Species richness
```{r}
library(gamm4)
# number of species
FV_MA_N_FL <- gamm4(N_species ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) + s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_fall) 

gam.check(FV_MA_N_FL$gam)
summary(FV_MA_N_FL$gam)
plot(FV_MA_N_FL$gam)


```


### Shannon-Weiner diversity
```{r}
# S-W
FV_MA_H_FL <- gamm4(H_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_fall) 

gam.check(FV_MA_H_FL$gam)
summary(FV_MA_H_FL$gam)
plot(FV_MA_H_FL$gam)

```

### Simpson's diveristy
```{r}

# Simpsons D
FV_MA_D_FL <- gamm4(D_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_fall) 

gam.check(FV_MA_D_FL$gam)
summary(FV_MA_D_FL$gam)
plot(FV_MA_D_FL$gam)

```

### Simpson's Evenness
```{r}
#simpsons e
FV_MA_E_FL <- gamm4(E_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) + s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_fall) 

gam.check(FV_MA_E_FL$gam)
summary(FV_MA_E_FL$gam)
plot(FV_MA_E_FL$gam)

```

### Taxonomic diversity
```{r}
#tax diversity
FV_MA_delta_FL <- gamm4(delta ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_fall) 

gam.check(FV_MA_delta_FL$gam)
summary(FV_MA_delta_FL$gam)
plot(FV_MA_delta_FL$gam)
```

### Taxonomic distinctness
```{r}
# tax distinctness
FV_MA_delta_star_FL <- gamm4(delta_star ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +   s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_fall)

gam.check(FV_MA_delta_star_FL$gam)
summary(FV_MA_delta_star_FL$gam)
plot(FV_MA_delta_star_FL$gam)

```

### Average taxonomic distinctness
```{r}
# avg tax distinctness
FV_MA_delta_plus_FL <- gamm4(delta_plus ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_fall) 

gam.check(FV_MA_delta_plus_FL$gam)
summary(FV_MA_delta_plus_FL$gam)
plot(FV_MA_delta_plus_FL$gam)

```

### Variation in taxonomic distinctness
```{r}
# var in tax distinctness
FV_MA_delta_var_FL <- gamm4(delta_var ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_fall) 

gam.check(FV_MA_delta_var_FL$gam)
summary(FV_MA_delta_var_FL$gam)
plot(FV_MA_delta_var_FL$gam)

```

## MADMF spring GAMMs
### Species richness
```{r}
# number of species
FV_MA_N_SP <- gamm4(N_species ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +   s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_spring) 

gam.check(FV_MA_N_SP$gam)
summary(FV_MA_N_SP$gam)
plot(FV_MA_N_SP$gam)

```

### Shannon-Weiner diversity
```{r}
# S-W
FV_MA_H_SP <- gamm4(H_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_spring) 

gam.check(FV_MA_H_SP$gam)
summary(FV_MA_H_SP$gam)
plot(FV_MA_N_SP$gam)
```

### Simpson's diversity
```{r}
# Simpsons D
FV_MA_D_SP <- gamm4(D_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_spring) 

gam.check(FV_MA_D_SP$gam)
summary(FV_MA_D_SP$gam)
plot(FV_MA_D_SP$gam)
```

### Simpson's Evenness
```{r}
#simpsons e
FV_MA_E_SP <- gamm4(E_index ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_spring) 

gam.check(FV_MA_E_SP$gam)
summary(FV_MA_E_SP$gam)
plot(FV_MA_E_SP$gam)
```

### Taxonomic diversity
```{r}
#tax diversity
FV_MA_delta_SP <- gamm4(delta ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_spring) 

gam.check(FV_MA_delta_SP$gam)
summary(FV_MA_delta_SP$gam)
plot(FV_MA_delta_SP$gam)
```


### Taxonomic distinctness
```{r}
# tax distinctness
FV_MA_delta_star_SP <- gamm4(delta_star ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_spring) 

gam.check(FV_MA_delta_star_SP$gam)
summary(FV_MA_delta_star_SP$gam)
plot(FV_MA_delta_star_SP$gam)
```

### Average taxonomic distinctness
```{r}
# avg tax distinctness
FV_MA_delta_plus_SP <- gamm4(delta_plus ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) + s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_spring) 

gam.check(FV_MA_delta_plus_SP$gam)
summary(FV_MA_delta_plus_SP$gam)
plot(FV_MA_delta_plus_SP$gam)
```

### Variation in taxonomic distinctness
```{r}
# var in tax distinctness
FV_MA_delta_var_SP <- gamm4(delta_var ~ s(FV_bot_temp) + s(FV_surf_temp) + s(FV_bot_sal) +  s(FV_surf_sal) +  s(DEPTH_M)+ s(LATITUDE, LONGITUDE) , random = ~ (1|YEAR), data = MA_spring) 

gam.check(FV_MA_delta_var_SP$gam)
summary(FV_MA_delta_var_SP$gam)
plot(FV_MA_delta_var_SP$gam)
```



