---
pagetitle: "Seagrant ME-NH trawl survey <br> Objectve 2: Nonmetric Multidimensional Scaling"
author: "Jerelle Jesse"
output: 
  html_document: 
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

```{r, echo = FALSE}
# Access GMRI CSS Style
library(gmRi)
gmRi::use_gmri_style_rmd(css_file = "gmri_rmarkdown.css")
```

# Nonmetric Multidimensional Scaling- Top 50 abundance species

## Data
* Top 50 abundant species
* average across depth strata using the [NOAA IEA technical document](https://noaa-edab.github.io/tech-doc/inshoresurvdat.html#data-analysis-29)

```{r }

#load packages
library(tidyverse)
library(vegan)
library(ggrepel)
library(ggforce)
library(ggnewscale)
library(ggthemes)
library(here)
library(plotly)
library(rmarkdown)

#set up ggplot theme
theme_set(theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
        axis.text.x = element_text(colour = "black", face = "bold", size = 12),
        legend.text = element_text(size = 12, face ="bold", colour ="black"),
        legend.position = "right", 
        axis.title.y = element_text(face = "bold", size = 14, angle=90),
        axis.title.x = element_text(face = "bold", size = 12, colour = "black"),
        legend.title = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
        legend.key=element_blank()))

#load in data
trawl_data_arrange<-read.csv(here("Data/species_abundance_matrix.csv"))[-1]
paged_table(head(trawl_data_arrange))

set.seed(123)
```

### Set up data for NMDS
* split community matrix into two dataframes- one for grouping variables and one for species biomass 
* calculate dissimilarity matrix with Bray-Curtis distances

```{r }
#set up final grouping data into dataframe
ME_group_data<-trawl_data_arrange[, c(1,2,3,55,56,57,58)]
ME_NMDS_data<-as.matrix(trawl_data_arrange[,4:53])

#calculate the distance matrix
ME_NMDS_distance<- vegdist(ME_NMDS_data, method="bray")
```

### Run the NMDS and extract scores
* change in community composition
* uses rank order
* stress < 0.2 is good, < 0.1 is great, <0.05 is excellent representation in reduced dimensions

```{r}
ME_NMDS=metaMDS(ME_NMDS_distance, # Our community-by-species matrix
                k=2, # The number of reduced dimensions
                method="bray",
                trymax=200) #increase iterations

#extract NMDS scores for ggplot
data.scores = as.data.frame(scores(ME_NMDS))
#add columns to data frame 
data.scores$Stratum = trawl_data_arrange$Stratum
data.scores$Region = trawl_data_arrange$Region
data.scores$Year = trawl_data_arrange$Year
data.scores$Season= trawl_data_arrange$Season
data.scores$Year_groups= trawl_data_arrange$YEAR_GROUPS
data.scores$Year_decades= trawl_data_arrange$YEAR_DECADES
data.scores$Region_new=trawl_data_arrange$REGION_NEW
data.scores$Region_year=trawl_data_arrange$REGION_YEAR
data.scores$Season_year=trawl_data_arrange$SEASON_YEAR

```

### Plots

#### Region
```{r }
# both seasons
data.scores_region<-filter(data.scores,Region_new!="Penobscot Bay")

#calculate center of ellipse for each region/year grouping
data.scores_2<-group_by(data.scores_region, Region_new, Year_groups)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Region_new, Year_groups, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:8)
rownames(d)<-NULL
centers<-full_join(d,id)


p <- ggplot() + 
  geom_point(data=data.scores_region, aes(x = NMDS1, y = NMDS2, shape=factor(Year_groups),color=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_region,aes(x=NMDS1,y=NMDS2,color=factor(Region_new)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_region, aes(x=NMDS1, y=NMDS2, color=Region_new),level = 0.95,size=1)+
  scale_color_colorblind(guide=FALSE)+
  labs(x = "NMDS1", y = "NMDS2")+
  geom_point(data=centers, aes(NMDS1, NMDS2,shape=Year_groups, color=Region_new), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2, color=Region_new), size=2)

#spring
data.scores_region<-filter(data.scores,Region_new!="Penobscot Bay")%>%
  filter(Season=="Spring")

#calculate center of ellipse for each region/year grouping
data.scores_2<-group_by(data.scores_region, Region_new, Year_groups, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Region_new, Year_groups,Season, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:8)
rownames(d)<-NULL
centers<-full_join(d,id)


p1 <- ggplot() + 
  geom_point(data=data.scores_region, aes(x = NMDS1, y = NMDS2, shape=factor(Year_groups),color=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_region,aes(x=NMDS1,y=NMDS2,color=factor(Region_new)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_region, aes(x=NMDS1, y=NMDS2, color=Region_new),level = 0.95,size=1)+
  scale_color_colorblind(guide=FALSE)+
  labs(x = "NMDS1", y = "NMDS2", title="Spring")+
  geom_point(data=centers, aes(NMDS1, NMDS2,shape=Year_groups, color=Region_new), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2, color=Region_new), size=2)

#Fall
data.scores_region<-filter(data.scores,Region_new!="Penobscot Bay")%>%
  filter(Season=="Fall")

#calculate center of ellipse for each region/year grouping
data.scores_2<-group_by(data.scores_region, Region_new, Year_groups, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Region_new, Year_groups,Season, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:8)
rownames(d)<-NULL
centers<-full_join(d,id)


p2 <- ggplot() + 
  geom_point(data=data.scores_region, aes(x = NMDS1, y = NMDS2, shape=factor(Year_groups),color=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_region,aes(x=NMDS1,y=NMDS2,color=factor(Region_new)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_region, aes(x=NMDS1, y=NMDS2, color=Region_new),level = 0.95,size=1)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Region", y = "NMDS2", shape = "Year",title="Fall")+
  geom_point(data=centers, aes(NMDS1, NMDS2,shape=Year_groups, color=Region_new), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2, color=Region_new), size=2)


#Spring
data.scores_east<-filter(data.scores, Region_new=="East of Penobscot Bay")%>%
  filter(Season=="Spring")
data.scores_west<-filter(data.scores, Region_new=="West of Penobscot Bay")%>%
  filter(Season=="Spring")

p3<- ggplot() + 
  geom_point(data=data.scores_east, aes(x = NMDS1, y = NMDS2,color=factor(Region_year), shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_east,aes(x=NMDS1,y=NMDS2,color=factor(Region_year)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_east, aes(x=NMDS1, y=NMDS2, color=Region_year),level = 0.95,size=1)+
  scale_color_manual(values=scales::seq_gradient_pal("grey", "black", "Lab")(seq(0,1,length.out=4)))+
  #geom_path(data=data.scores_east, aes(x=NMDS1,y=NMDS2,group=factor(Region_new)))+
  #geom_text_repel(data=data.scores_east,aes(x=NMDS1,y=NMDS2,label=Year), size=4, fontface="bold", color="black")+
  ggnewscale::new_scale_color()+
  geom_point(data=data.scores_west, aes(x = NMDS1, y = NMDS2,color=factor(Region_year), shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_west,aes(x=NMDS1,y=NMDS2,color=factor(Region_year)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_west, aes(x=NMDS1, y=NMDS2, color=Region_year),level = 0.95,size=1)+
  scale_color_manual(values=scales::seq_gradient_pal("light blue", "navy", "Lab")(seq(0,1,length.out=4)))+
  #geom_path(data=data.scores_west, aes(x=NMDS1,y=NMDS2,group=factor(Region_new)))+
  #geom_text_repel(data=data.scores_west,aes(x=NMDS1,y=NMDS2,label=Year), size=4, fontface="bold", color="black")+
  labs(x = "NMDS1", colour = "Region/Year", y = "NMDS2", shape = "Region", fill="Region/Year", title="Spring")


#Fall
data.scores_east<-filter(data.scores, Region_new=="East of Penobscot Bay")%>%
  filter(Season=="Fall")
data.scores_west<-filter(data.scores, Region_new=="West of Penobscot Bay")%>%
  filter(Season=="Fall")

p4<- ggplot() + 
  geom_point(data=data.scores_east, aes(x = NMDS1, y = NMDS2,color=factor(Region_year), shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_east,aes(x=NMDS1,y=NMDS2,color=factor(Region_year)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_east, aes(x=NMDS1, y=NMDS2, color=Region_year),level = 0.95,size=1)+
  scale_color_manual(values=scales::seq_gradient_pal("grey", "black", "Lab")(seq(0,1,length.out=4)))+
  #geom_path(data=data.scores_east, aes(x=NMDS1,y=NMDS2,group=factor(Region_new)))+
  #geom_text_repel(data=data.scores_east,aes(x=NMDS1,y=NMDS2,label=Year), size=4, fontface="bold",   color="black")+
  ggnewscale::new_scale_color()+
  geom_point(data=data.scores_west, aes(x = NMDS1, y = NMDS2,color=factor(Region_year),      shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_west,aes(x=NMDS1,y=NMDS2,color=factor(Region_year)), size=1,   alpha=0.15)+
  stat_ellipse(data=data.scores_west, aes(x=NMDS1, y=NMDS2, color=Region_year),level = 0.95,size=1)+
  scale_color_manual(values=scales::seq_gradient_pal("light blue", "navy", "Lab")(seq(0,1,length.out=4)))+
  #geom_path(data=data.scores_west, aes(x=NMDS1,y=NMDS2,group=factor(Region_new)))+
  #geom_text_repel(data=data.scores_west,aes(x=NMDS1,y=NMDS2,label=Year), size=4, fontface="bold",   color="black")+
  labs(x = "NMDS1", colour = "Region/Year", y = "NMDS2", shape = "Year", fill="Region/Year", title="Fall")

p
p1
p2
p3
p4

```

#### Time
```{r }
#both seasons
data.scores_region<-filter(data.scores,Region_new!="Penobscot Bay")

#centers for year plots
data.scores_2<-group_by(data.scores_region, Year_groups)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Year_groups, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:4)
rownames(d)<-NULL
centers<-full_join(d,id)


p <- ggplot() + 
  geom_point(data=data.scores_region, aes(x = NMDS1, y = NMDS2,color=factor(Year_groups), shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_region,aes(x=NMDS1,y=NMDS2,color=factor(Year_groups)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_region, aes(x=NMDS1, y=NMDS2, color=Year_groups),level = 0.95,size=1)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Year", y = "NMDS2", shape = "Region")+
  geom_point(data=centers, aes(NMDS1, NMDS2, color=Year_groups), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2), size=2)

#spring
data.scores_region<-filter(data.scores,Region_new!="Penobscot Bay")%>%
  filter(Season=="Spring")
#centers for year plots
data.scores_2<-group_by(data.scores_region, Year_groups, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Year_groups,Season, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:4)
rownames(d)<-NULL
centers<-full_join(d,id)


p1 <- ggplot() + 
  geom_point(data=data.scores_region, aes(x = NMDS1, y = NMDS2,color=factor(Year_groups), shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_region,aes(x=NMDS1,y=NMDS2,color=factor(Year_groups)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_region, aes(x=NMDS1, y=NMDS2, color=Year_groups),level = 0.95,size=1)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Year", y = "NMDS2", shape = "Region", title="Spring")+
  geom_point(data=centers, aes(NMDS1, NMDS2, color=Year_groups), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2), size=2)


#Fall
data.scores_region<-filter(data.scores,Region_new!="Penobscot Bay")%>%
  filter(Season=="Fall")
#centers for year plots
data.scores_2<-group_by(data.scores_region, Year_groups, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Year_groups,Season, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:4)
rownames(d)<-NULL
centers<-full_join(d,id)

p2 <- ggplot() + 
  geom_point(data=data.scores_region, aes(x = NMDS1, y = NMDS2,color=factor(Year_groups), shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_region,aes(x=NMDS1,y=NMDS2,color=factor(Year_groups)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_region, aes(x=NMDS1, y=NMDS2, color=Year_groups),level = 0.95,size=1)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Year", y = "NMDS2", shape = "Region", title="Fall")+
  geom_point(data=centers, aes(NMDS1, NMDS2, color=Year_groups), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2), size=2)




#Spring
data.scores_region<-filter(data.scores,Region_new!="Penobscot Bay")%>%
  filter(Season=="Spring")

#centers for year plots
data.scores_2<-group_by(data.scores_region, Year_groups,Region_new, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Year_groups,Season,Region_new, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:8)
rownames(d)<-NULL
centers<-full_join(d,id)

p3 <- ggplot() + 
  geom_point(data=data.scores_region, aes(x = NMDS1, y = NMDS2,color=factor(Year_groups), shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_region,aes(x=NMDS1,y=NMDS2,color=factor(Year_groups)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_region, aes(x=NMDS1, y=NMDS2, color=Year_groups),level = 0.95,size=1)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Year", y = "NMDS2", shape = "Region", title="Spring")+
  geom_point(data=centers, aes(NMDS1, NMDS2, color=Year_groups, shape=Region_new), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2, group=Region_new), size=2)


#Fall
data.scores_region<-filter(data.scores,Region_new!="Penobscot Bay")%>%
  filter(Season=="Fall")

#centers for year plots
data.scores_2<-group_by(data.scores_region, Year_groups,Region_new, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Year_groups,Season,Region_new, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:8)
rownames(d)<-NULL
centers<-full_join(d,id)

p4 <- ggplot() + 
  geom_point(data=data.scores_region, aes(x = NMDS1, y = NMDS2,color=factor(Year_groups), shape=factor(Region_new)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_region,aes(x=NMDS1,y=NMDS2,color=factor(Year_groups)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_region, aes(x=NMDS1, y=NMDS2, color=Year_groups),level = 0.95,size=1)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Year", y = "NMDS2", shape = "Region", title="Fall")+
  geom_point(data=centers, aes(NMDS1, NMDS2, color=Year_groups, shape=Region_new), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2, group=Region_new), size=2)


p
p1
p2
p3
p4

```

#### Season
```{r }
#all regions
#centers for year plots
data.scores_2<-group_by(data.scores, Year_groups, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Year_groups,Season, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
  dplyr:: select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:8)
rownames(d)<-NULL
centers<-full_join(d,id)


p <- ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) + 
  geom_point(size = 3, aes(shape=factor(Year_groups), color=factor(Season)))+ 
  stat_ellipse(data=data.scores, aes(x=NMDS1, y=NMDS2, color=Season),level = 0.95,size=1)+
  #geom_mark_ellipse(aes(color=factor(Season)), size=1, alpha=0.15)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Season", y = "NMDS2", shape = "Year", fill="Season")+
    geom_point(data=centers, aes(NMDS1, NMDS2, color=Season, shape=Year_groups), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2, color=Season), size=2)


#ellipse season and year plot

data.scores_west<-filter(data.scores, Region_new=="West of Penobscot Bay")
#centers for year plots
data.scores_2<-group_by(data.scores_west, Year_groups,Region_new, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Year_groups,Season,Region_new, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
  dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:8)
rownames(d)<-NULL
centers<-full_join(d,id)

#West
p1 <- ggplot(data.scores_west, aes(x = NMDS1, y = NMDS2)) + 
  geom_point(size = 3, aes(shape=factor(Year_groups), color=factor(Season)))+ 
  stat_ellipse(data=data.scores_west, aes(x=NMDS1, y=NMDS2, color=Season),level = 0.95,size=1)+
  #geom_mark_ellipse(aes(color=factor(Season)), size=1, alpha=0.15)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Season", y = "NMDS2", shape = "Year", fill="Season", title="West")+
    geom_point(data=centers, aes(NMDS1, NMDS2, color=Season, shape=Year_groups), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2, color=Season), size=2)

#East
data.scores_east<-filter(data.scores, Region_new=="East of Penobscot Bay")
#centers for year plots
data.scores_2<-group_by(data.scores_east, Year_groups,Region_new, Season)%>%
   mutate(indicator=cur_group_id())
id<-dplyr::select(data.scores_2, Year_groups,Season,Region_new, indicator)
id<-unique(id)

ctr<-NULL
Ncombo<-length(unique(data.scores_2$indicator))

for (i in 1:Ncombo) {
 combos<- filter(data.scores_2, indicator==i)%>%
   ungroup()%>%
   dplyr::select(1:2)
 
  ctr[[i]]<-MASS::cov.trob(combos)$center
  
}

d<-as.data.frame(ctr)
d<-as.data.frame(t(d))
d$indicator<-seq(1:8)
rownames(d)<-NULL
centers<-full_join(d,id)


p2 <- ggplot(data.scores_east, aes(x = NMDS1, y = NMDS2)) + 
  geom_point(size = 3, aes(shape=factor(Year_groups), color=factor(Season)))+ 
  stat_ellipse(data=data.scores_east, aes(x=NMDS1, y=NMDS2, color=Season),level = 0.95,size=1)+
  #geom_mark_ellipse(aes(color=factor(Season)), size=1, alpha=0.15)+
  scale_color_colorblind()+
  labs(x = "NMDS1", colour = "Season", y = "NMDS2", shape = "Year", fill="Season", title="East")+
  geom_point(data=centers, aes(NMDS1, NMDS2, color=Season, shape=Year_groups), size=10)+
  geom_path(data=centers, aes(NMDS1, NMDS2, group=Season), size=2)

#West
data.scores_spring<-filter(data.scores, Season=="Spring")%>%
  filter(Region_new=="West of Penobscot Bay")
data.scores_fall<-filter(data.scores, Season=="Fall")%>%
  filter(Region_new=="West of Penobscot Bay")

p3 <- ggplot() + 
  geom_point(data=data.scores_fall, aes(x = NMDS1, y = NMDS2,color=factor(Season_year), shape=factor(Season)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_fall,aes(x=NMDS1,y=NMDS2,color=factor(Season_year)), size=1, alpha=0.15)+
stat_ellipse(data=data.scores_fall, aes(x=NMDS1, y=NMDS2, color=Season_year),level = 0.95,size=1)+
  scale_color_manual(values=scales::seq_gradient_pal("grey", "black", "Lab")(seq(0,1,length.out=4)))+
  #geom_path(data=data.scores_fall, aes(x=NMDS1,y=NMDS2,group=factor(Season)))+
  #geom_text_repel(data=data.scores_fall,aes(x=NMDS1,y=NMDS2,label=Year), size=4, fontface="bold", color="black")+
  ggnewscale::new_scale_color()+
  geom_point(data=data.scores_spring, aes(x = NMDS1, y = NMDS2,color=factor(Season_year), shape=factor(Season)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_spring,aes(x=NMDS1,y=NMDS2,color=factor(Season_year)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_spring, aes(x=NMDS1, y=NMDS2, color=Season_year),level = 0.95,size=1)+
  scale_color_manual(values=scales::seq_gradient_pal("light blue", "navy", "Lab")(seq(0,1,length.out=4)))+
  #geom_path(data=data.scores_spring, aes(x=NMDS1,y=NMDS2,group=factor(Season)),color="navy")+
  #geom_text_repel(data=data.scores_spring,aes(x=NMDS1,y=NMDS2,label=Year), size=4, fontface="bold", color="black")+
  labs(x = "NMDS1", colour = "Season/Year", y = "NMDS2", shape = "Season", fill="Season/Year", title="West")


#East
data.scores_spring<-filter(data.scores, Season=="Spring")%>%
  filter(Region_new=="East of Penobscot Bay")
data.scores_fall<-filter(data.scores, Season=="Fall")%>%
  filter(Region_new=="East of Penobscot Bay")

p4 <- ggplot() + 
  geom_point(data=data.scores_fall, aes(x = NMDS1, y = NMDS2,color=factor(Season_year), shape=factor(Season)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_fall,aes(x=NMDS1,y=NMDS2,color=factor(Season_year)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_fall, aes(x=NMDS1, y=NMDS2, color=Season_year),level = 0.95,size=1)+
  scale_color_manual(values=scales::seq_gradient_pal("grey", "black", "Lab")(seq(0,1,length.out=4)))+
  #geom_path(data=data.scores_fall, aes(x=NMDS1,y=NMDS2,group=factor(Season)))+
  #geom_text_repel(data=data.scores_fall,aes(x=NMDS1,y=NMDS2,label=Year), size=4, fontface="bold", color="black")+
  ggnewscale::new_scale_color()+
  geom_point(data=data.scores_spring, aes(x = NMDS1, y = NMDS2,color=factor(Season_year), shape=factor(Season)),size = 3)+ 
  #geom_mark_ellipse(data=data.scores_spring,aes(x=NMDS1,y=NMDS2,color=factor(Season_year)), size=1, alpha=0.15)+
  stat_ellipse(data=data.scores_spring, aes(x=NMDS1, y=NMDS2, color=Season_year),level = 0.95,size=1)+
  scale_color_manual(values=scales::seq_gradient_pal("light blue", "navy", "Lab")(seq(0,1,length.out=4)))+
  #geom_path(data=data.scores_spring, aes(x=NMDS1,y=NMDS2,group=factor(Season)),color="navy")+
  #geom_text_repel(data=data.scores_spring,aes(x=NMDS1,y=NMDS2,label=Year), size=4, fontface="bold", color="black")+
  labs(x = "NMDS1", colour = "Season/Year", y = "NMDS2", shape = "Year", fill="Season/Year", title="East")

p
p1
p2
p3
p4
```

